\section{Panel Overview}

The weekly cryptocurrency panel used in this thesis is assembled from a set of
public data sources and aligned to a common weekly (week-ending) calendar.
The base panel, \path{crypto_panel_features.npz}, has $T = 324$ calendar weeks
(2020-01-05 to 2026-03-15), $N = 86$ tickers, one next-week return target, and
49 predictor columns.
The augmented panel, \path{crypto_panel_features_with_etf.npz}, appends four
granular Farside ETF-flow variables, giving 53 predictor columns in total.
The six information sets used in the empirical comparisons draw on a 42-feature
subset of these columns; the remaining 11 columns are DeFi and derivatives
features that are retained in the stored panel but excluded from the model
comparisons in this thesis.

\Cref{tab:data_sources} documents the provider, access method, and coverage for
every feature group, so that the panel can in principle be reconstructed from
its original sources rather than only from the saved arrays.
All primary sources are public and most require no API key; the only manual
download step is the Farside ETF flow table, which is distributed as a CSV.

Missing values are encoded as $-99.99$ and masked in both loss computation and
portfolio evaluation.
Features that begin only partway through the sample---most importantly the spot
ETF flows (Bitcoin ETFs launched 2024-01-11, Ethereum ETFs 2024-07-23) and the
Trump social-media series---are coded as missing before they exist rather than
back-filled, so that no future information is introduced.
The fixed 86-token universe (the largest tokens by market capitalization,
excluding stablecoins and wrapped assets) is best interpreted as the asset
universe selected near the end of the sample, not as a fully auditable
historical listing universe; we return to the resulting survivorship caveat in
\Cref{sec:data:survival}.

\section{Data Sources and Acquisition}
\label{sec:data:sources}

\Cref{tab:data_sources} lists the upstream provider and download method for
each feature group.
Each source is queried over the full sample where coverage permits, and daily
or post-level observations are aggregated to the panel's weekly (week-ending)
calendar.

\begin{table}[!b]
\centering
\caption{Data sources and acquisition by feature group. All sources are public;
         only the Farside ETF tables require a manual CSV download. Daily and
         post-level series are aggregated to the weekly panel calendar.}
\label{tab:data_sources}
\small
\begin{tabularx}{\linewidth}{>{\raggedright\arraybackslash}p{0.22\linewidth} >{\raggedright\arraybackslash}p{0.20\linewidth} X}
\toprule
\thead{Feature group (count)} & \thead{Provider} & \thead{Access method and coverage} \\
\midrule
Price \& momentum (5), Technical (6)
  & Yahoo Finance \citep{yahoo_finance}
  & \texttt{yfinance} library, no key; \texttt{\{SYMBOL\}-USD} weekly OHLCV for
    the 86-token universe, from which returns and technical indicators are
    computed. Full sample. \\
On-chain (5)
  & Coin Metrics \citep{coinmetrics}
  & Community REST API, no key; active addresses, transaction count, NVT, MVRV,
    and exchange net flow. Per asset where the metric is available. \\
Macro \& sentiment (11)
  & alternative.me; Yahoo Finance \citep{alternativeme_fng,yahoo_finance}
  & Fear \& Greed index, plus S\&P~500, DXY, VIX, gold (GLD), silver (SLV), and
    Dow (DIA) returns and relative volume. Common to all assets. \\
Aggregate ETF flow \& volume (5)
  & Farside; Yahoo Finance \citep{bitcoin_etf_farside,ethereum_etf_farside,yahoo_finance}
  & Farside CSV (manual download) plus \texttt{yfinance} aggregate U.S. spot ETF
    volume. BTC ETFs from 2024-01-11, ETH ETFs from 2024-07-23. \\
Prediction market (1)
  & Polymarket \citep{polymarket}
  & Gamma and CLOB APIs, no key; volume-weighted BTC price-target probability,
    logit-transformed. \\
Granular ETF flow (4)
  & Farside Investors \citep{bitcoin_etf_farside,ethereum_etf_farside}
  & Farside CSV (manual download); GBTC, BTC ex-GBTC, ETHE, and ETH ex-ETHE
    weekly flows. \\
Trump social media (5)
  & Truth Social / X archives
  & Post archives: Truth Social (2022-02 to 2024-06) and X/Twitter (2025-01 to
    2026-03); weekly post count, capitalization ratio, tariff and crypto topic
    scores, and sentiment. Weeks outside coverage are set to missing. \\
\midrule
Market cap (weighting)
  & CoinGecko \citep{coingecko}
  & Free API; weekly market capitalization used for price-weighting and the
    market benchmark, not as a predictor. \\
DeFi \& derivatives (11, unused)
  & DefiLlama; Binance Futures
  & TVL, DEX volume, fees, and stablecoin cap, plus futures funding rate, open
    interest, and long/short ratio. Stored in the panel but excluded from the
    42-feature comparisons. \\
\bottomrule
\end{tabularx}
\end{table}

Each weekly series is normalized in one of two ways, depending on whether it
varies across assets.
Per-asset features (price/momentum, technical, and on-chain) are
cross-sectionally rank-normalized within each week, which removes cross-week
scale dependence and introduces no look-ahead.
Market-wide series (macro and sentiment, aggregate ETF flow and volume, the
Polymarket probability, the granular Farside flows, and the Trump features)
share the same value across assets in a given week and are standardized with a
trailing 52-week rolling z-score (minimum four observations), clipped to
$[-3, 3]$.

For the spot ETF flows, the Farside Investors tables
\citep{bitcoin_etf_farside,ethereum_etf_farside} are downloaded as CSV files and
parsed so that parenthesized values denote outflows (negative), dashes denote
missing or zero-flow cells, and daily rows are summed over the interval from the
previous panel date to the current panel date.
The four granular variables appended in the final panel are GBTC flow, total
Bitcoin ETF flow excluding GBTC, ETHE flow, and total Ethereum ETF flow
excluding ETHE, each then standardized with the same 52-week rule.

\section{Feature Categories}

\Cref{tab:features} summarizes the 42 features actually used in the model
comparisons.
The table reports variable groups and representative examples; the upstream
provider for each group is given in \Cref{tab:data_sources}.
The grouping reflects the main conceptual distinction in the paper: some
signals are slow-moving and structural, whereas others are event-like and
time-dependent.

\begin{table}[t]
\centering
\caption{Feature categories used in the empirical model comparisons.
         $M_g$ = number of features in group $g$.}
\label{tab:features}
\small
\begin{tabular}{llc}
\toprule
\thead{Group} & \thead{Examples} & \thead{$M_g$} \\
\midrule
Price \& Momentum    & 1w/4w/12w/26w/52w returns               & 5  \\
Technical            & RSI, Bollinger \%, ATR, OBV, volume ratio & 6  \\
On-chain             & Active addresses, NVT, MVRV, net flow   & 5  \\
Macro \& Sentiment   & Fear-greed index, equity, FX, volatility, metal returns & 11 \\
ETF \& Prediction Market & BTC/ETH ETF flow fields, BTC probability field & 6  \\
Trump Social Media   & Post count, caps ratio, tariff score, crypto score, sentiment & 5  \\
Farside ETF Granular & GBTC, BTC ex-GBTC, ETHE, ETH ex-ETHE    & 4  \\
\midrule
Total                &                                          & 42 \\
\bottomrule
\end{tabular}
\end{table}

\section{Data Splits}
\label{sec:data:splits}

We partition the panel chronologically into training (70\%, weeks 0--226),
validation (15\%, weeks 227--274), and test (15\%, weeks 275--323),
corresponding to 49 held-out panel weeks.
Because the saved prediction masks exclude one test week from valid
long-short portfolio construction, the reported Sharpe ratios are computed
from 48 weekly portfolio-return observations.
Hyperparameters are selected on the validation set, and the test set is used
only for final reporting.
This preserves the temporal ordering of the problem and avoids contaminating
evaluation with repeated test-period feedback.

\section{Look-ahead Bias}
\label{sec:data:lookahead}

The replication code applies time-series standardization using a rolling
52-week z-score.
At week $t$, the mean and standard deviation are computed over information
available through week $t$.
No future panel dates enter the normalization window.

For the Farside ETF variables, daily rows are first assigned to the
corresponding panel week and then standardized.
The inclusion of week $t$ in its own rolling mean gives the current
observation weight $1/52 \approx 2\%$ once the full window is available.
This avoids future leakage, although it still means that weekly ETF-flow
features are contemporaneous weekly aggregates rather than signals observed
strictly before the start of the week.
The same timing interpretation applies to all market-wide weekly predictors
that are standardized with the rolling z-score rather than rank-normalized
within each week.

\section{Provenance and Survivorship Limitations}
\label{sec:data:survival}

Two limitations should be kept in mind.
First, several primary sources are community or free-tier endpoints
(CoinMetrics Community, the CoinGecko free API, alternative.me, and Polymarket)
whose historical values can be revised or rate-limited, and the Trump
social-media archive has coverage gaps that are coded as missing.
These affect the exact reproducibility of individual cells, not the internal
consistency of the model comparisons: every method uses the same panel, feature
configurations, masks, and chronological splits.

Second, the fixed 86-token universe may contain survivorship bias.
The repository provides the final ticker list but not a complete historical
listing-and-delisting rule.
Tokens that were historically relevant but absent from the fixed universe
would therefore be excluded from the analysis.
The paper should consequently be read primarily as evidence on the relative
suitability of objective functions and model architectures for a fixed
cross-sectional cryptocurrency ranking problem, rather than as an
unconditional claim about live investability of the entire cryptocurrency
market.
